Nuclei Segmentation in Histopathology images using Deep Neural Networks

### 2016-11-07 Paper.tex
tiny paragraph

In \cite{Drelie08-298}, they have a partially annotated data set of 
histopathology images for nuclei segmentation.

@@

%% Contribution as a dataset, small introduction
One of the main contributions of this paper is the now public available 
nuclei detection dataset within HE stainned histopathology images which 
can be found at website \url{http://cbio.mines-paristech.fr/~pnaylor/Downloads/HECellSegmentation.zip} 
(not a true address). This annotated dataset provides images clustered by 
patients. Each patient has at least 3 annotated $512 \times 512$ HE 
stainned histopathology images with their associated ground truth. Each 
ground truth image is a $512 \times 512$ image where each pixel value above 
$0$ is considered as a pixel belonging to a nuclei. The differences in 
values of these pixels denote different nucleus, such an annotation can 
be used in several processing step which does or doesn't take into 
account clutered nuclei. See figure \ref{fig:annotation} for an example of 
three annotated images. This annotation was conducted via the help of 
the software ITK-snap, \cite{py06nimg} \url{http://www.itksnap.org}, and were annotated by the authors of the paper.

@@

\section{Methodology}
\label{sec:method}
%% Segmentation task
Let A be the space of RGB images, A can typically be 
$\mathbb{R}^{n \times p \times 3}$ and let B the space B the space of annotation 
images, in our case $\{0,1\}^{n \times p}$. We have a set of 
$(A_l,B_l)_{l \in [1, N]}$ for a supervised learning approach. Our goal is 
a prediction task 
named as semantic segmentation, we wish to maximize the prediction of 
an unseen element belonging to B given an new element in A. We maximize 
thus prediction by modelizing our prediction function as the softmax output of a deep neural 
network. We find the model parameters by minimizing a log loss function 
defined as:
 $\frac{1}{\sum_{i,j}w_{ij}} \sum_{i,j} \sum_k w_{i,j} t_{i,j,k} \log (\widehat{p_{i,j,k}})$
, where $k$ designates a certain label, $w_{i,j}$ is a 
certain weight given to pixel $i,j$, $t_{i,j,k}$ is equal to 1 if pixel $i,j$ is 
of class $k$ and $\widehat{p_{i,j,k}}$ designates the estimated probability 
of pixel $i,j$ of being $k$ via the softmax output of the neural network.
We minimize the loss function via stochastic gradient descent.

@@

\section{Different architectures/Results}
\label{sec:results}
We experience with 3 known arhitectures in semantic segmentation, named PangNet, Fully Convolutionnal Net (FCN) and DeconvNet. The most basic architecture, PangNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map, for more information please refer to \cite{pang2010cell}. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture, for a more detailed explaination please refer to \cite{long2015fcn}. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned, for more information please 
refer to the original paper \cite{noh2015learning}. We did not
ensemble classifier as suggested in 
\cite{noh2015learning} as our FCN and DeconvNet classifier achieve very similar scores on the test sets.
% The final architecture that we tried is the UNet. This network combines 
%both contribution of the above networks, the UNet has a similar 
%architecture to the DeconvNet with skip layers.
\section{Result and discussion}
\label{sec:result}
We conducted our experiments on GPUs via the caffe framework, 
\cite{jia2014caffe}. We present our results in table \ref{tab:res} where 
we averaged the metrics other all left out patients. We also added, for 
illustration purposes \ref{fig:prediction}, image predictions and 
associated 
probability maps to evaluate some differences between the classifiers. 
We notice that the simpler net, PangNet seems to have learnt simple 
rules such as color information. As we can not notice by the "holes" in 
the probability map when the interior of the cell is very bright. On the 
contrary, deeper nets as FCN and DeconvNet have learnt deeper features 
and seem capable of recognizing whole cells. On these probability maps 
we applied our watershed segmentation and we can picture the results in 
the third 
column. Our post-processing 
scheme is very sensitive to noise as we can see in the PangNet 
prediction, many unrelated minimas will lead to highly partitionned cell. 
This phenomenom also appears with the DeconvNet where undesired 
cuts were performed, like for the cell at the bottom left of the image. 
We picked $\lambda = 7$ which seemed to give a nice partitionning of 
the image, this value can be found by cross validation or by simple 
visualation of the results.
\begin{table}
\begin{tabular}{|c|c|c|c|}
\hline
  & PangNet & DeconvNet & FCN \\
 \hline
Accuracy  &       0.936 & 0.968 &0.958  \\
IU   &    0.759 &     0.856 & 0.857 \\
Recall     &       0.744  &     0.856 & 0.855 \\
Precision   &       0.741 &     0.858 & 0.878\\
F1    &       0.742 &     0.856 & 0.866 \\
TP       &       0.744 &     0.856 & 0.855  \\
TN      &       0.963 &     0.981 &  0.977\\
Performance    &       0.853 &     0.918 & 0.916\\
Pixel error  &  0.063 &     0.032 & 0.042\\
\hline
\end{tabular}
\caption{Results}
\label{tab:res}
\end{table}

>>>>

\noindent In \cite{Drelie08-298}, they have a partially annotated data set for segmentation of 
histopathology images for nuclei segmentation.   However their study does not sollely focus on tripple negativ breast cancer patients and not only segmentation.

@@

%% Contribution as a dataset, small introduction
\noindent One of the main contributions of this paper is the now public available 
nuclei detection dataset within HE stainned histopathology images which 
can be found at website \url{http://cbio.mines-paristech.fr/~pnaylor/Downloads/HECellSegmentation.zip} 
(not a true address). This annotated dataset provides images clustered by 
patients. Each patient has at least 3 annotated $512 \times 512$ HE 
stainned histopathology images with their associated ground truth. Each 
ground truth image is a $512 \times 512$ image where each pixel value above 
$0$ is considered as a pixel belonging to a nuclei. The differences in 
values of these pixels denote different nucleus, such an annotation can 
be used in several processing step which does or doesn't take into 
account clutered nuclei. See figure \ref{fig:annotation} for an example of 
three annotated images. This annotation was conducted via the help of 
the software ITK-snap, \cite{py06nimg} \url{http://www.itksnap.org}, and were annotated by the authors of the paper.

@@

\section{Methodology}
\label{sec:method}
%% Segmentation task
\noindent Let A be the space of RGB images, A can typically be 
$\mathbb{R}^{n \times p \times 3}$ and let B the space B the space of annotation 
images, in our case $\{0,1\}^{n \times p}$. We have a set of 
$(A_l,B_l)_{l \in [1, N]}$ for a supervised learning approach. Our goal is 
a prediction task 
named as semantic segmentation, we wish to maximize the prediction of 
an unseen element belonging to B given an new element in A. We maximize 
thus prediction by modelizing our prediction function as the softmax output of a deep neural 
network. We find the model parameters by minimizing a log loss function 
defined as:
 $\frac{1}{\sum_{i,j}w_{ij}} \sum_{i,j} \sum_k w_{i,j} t_{i,j,k} \log (\widehat{p_{i,j,k}})$
, where $k$ designates a certain label, $w_{i,j}$ is a 
certain weight given to pixel $i,j$, $t_{i,j,k}$ is equal to 1 if pixel $i,j$ is 
of class $k$ and $\widehat{p_{i,j,k}}$ designates the estimated probability 
of pixel $i,j$ of being $k$ via the softmax output of the neural network.
We minimize the loss function via stochastic gradient descent.

@@

\section{Different architectures/Results}
\label{sec:results}
We experience with 3 known arhitectures in semantic segmentation, named PangNet, Fully Convolutionnal Net (FCN) and DeconvNet. The most basic architecture, PangNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map, for more information please refer to \cite{pang2010cell}. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture, for a more detailed explaination please refer to \cite{long2015fcn}. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned, for more information please 
refer to the original paper \cite{noh2015learning}. We did not
ensemble classifier as suggested in 
\cite{noh2015learning} as our FCN and DeconvNet classifier achieve very similar scores on the test sets.
% The final architecture that we tried is the UNet. This network combines 
%both contribution of the above networks, the UNet has a similar 
%architecture to the DeconvNet with skip layers.
\section{Result and discussion}
\label{sec:result}
\noindent We conducted our experiments on GPUs via the caffe framework, 
\cite{jia2014caffe}. We present our results in table \ref{tab:res} where 
we averaged the metrics other all left out patients. We also added, for 
illustration purposes \ref{fig:prediction}, image predictions and 
associated 
probability maps to evaluate some differences between the classifiers. 
We notice that the simpler net, PangNet seems to have learnt simple 
rules such as color information. As we can not notice by the "holes" in 
the probability map when the interior of the cell is very bright. On the 
contrary, deeper nets as FCN and DeconvNet have learnt deeper features 
and seem capable of recognizing whole cells. On these probability maps 
we applied our watershed segmentation and we can picture the results in 
the third 
column. Our post-processing 
scheme is very sensitive to noise as we can see in the PangNet 
prediction, many unrelated minimas will lead to highly partitionned cell. 
This phenomenom also appears with the DeconvNet where undesired 
cuts were performed, like for the cell at the bottom left of the image. 
We picked $\lambda = 7$ which seemed to give a nice partitionning of 
the image, this value can be found by cross validation or by simple 
visualation of the results.
\begin{table}
\begin{tabular}{|c|c|c|c|}
\hline
  & PangNet & DeconvNet & FCN \\
 \hline
Accuracy  &       0.936 & 0.968 &0.958  \\
IU   &    0.759 &     0.856 & 0.857 \\
Recall     &       0.744  &     0.856 & 0.855 \\
Precision   &       0.741 &     0.858 & 0.878\\
F1    &       0.742 &     0.856 & 0.866 \\
TP       &       0.744 &     0.856 & 0.855  \\
TN      &       0.963 &     0.981 &  0.977\\
Performance    &       0.853 &     0.918 & 0.916\\
Pixel error  &  0.063 &     0.032 & 0.042\\
\hline
\end{tabular}
\caption{Results}
\label{tab:res}
\end{table}
